---
title: "New York Times Web API"
author: "Nnaemeka Okereafor"
date: "10/24/2021"
output: rmdformats::readthedown 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(jsonlite)
library(httr)
library(kableExtra)
```

### About the Assignment

In the Assignment I created an API key for New York Times database to read articles. Amongst the numerous articles, I chose to request for the most viewed articles for the last seven days.The response was received as a JSON data which I converted into a dataframe.

### Approach

I signed up with the New York Times and created an API key to enable me send a request to the New York Times server through the *GET()* function of the *httr* package and the response was saved in an object named *article*. The object was displayed to view its status, content-type, date and etc. The status was 200 and the content-type was in application/json format. The status and content information signified that there was a connection for the server to send the json formatted request. I displayed the header to have an overview of the *article* object. Then the unicode formatted content of the object was converted to a character vector with the *content()* function parsed with *as* argument equals *text* before parsing it the *fromJSON()* function. The *fromJSON()* function converts the json formatted content into a dataframe of articles, copyright, number of articles, and status.


### Send Request to get most viewed articles for the last seven days and save the response in a variable

```{r pressure, echo=TRUE}
url_viewed <-"https://api.nytimes.com/svc/mostpopular/v2/viewed/1.json?api-key=jsWenA4LsyGDPE3LOkNnt4kxlQbG1clD"
article <- GET(url_viewed)
```

### Display response

```{r}
article
```
 
 
```{r}
article_header <- article$headers
article_header

```

### View the unicode

```{r}
str(article$content)
```

### Convert the unicode to character vector

```{r}
nyt_article<-content(article, as="text")
#nyt_article
```

### Parse the character vector into a fromJSON to convert it to a list

```{r}
nyt_articlent_json<-nyt_article %>% fromJSON
#glimpse(nyt_articlent_json)
```

The glimpse() was called to unbundle the content of the nyt_articlent_json list to view its elements. The list contains four elements, namely-status, copyright, num_results and results dataframe. 


### status

```{r}
nyt_articlent_json$status

```

### copyright

```{r}
nyt_articlent_json$copyright

```

### Number of observations

```{r}
nyt_articlent_json$num_results

```

### Articles

```{r}
samp_article<-nyt_articlent_json$results

kbl(samp_article)%>%
  kable_styling() %>% kable_paper("hover", full_width = F)

```


### Conclusion

It is a nice assignment for students of data science in the age of globalization. I learnt a lot in the assignment and I recommend that anyone in quest to learn data science should avail himself to learn about APIs. Now, I have the opportunity to view the most viewed article of any week in the comfort of my home with a click of my computer mouse.



