---
title: "Tidying and Exploratory Analysis of Indian stock Data"
author: "Nnaemeka Okereafor"
date: "10/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
library(tidyverse)
```


### About The Project

The dataset contains data from a list of Indian stocks in NSE. It includes a collection of well performing stocks. Per Jiho Kim,The dataset contains lots of #NA and #REF observations. Find the trading high price mean and low price mean per sector and the distribution of the mean prices. Dataset source: [link](https://www.kaggle.com/souravroy1/stock-market-data/download) 

**Read dataset from Github**

```{r}
stock <- read.csv(file="https://raw.githubusercontent.com/nnaemeka-git/global-datasets/main/stock%20dataset.csv", sep=",")
as_tibble(stock)
```

**Remove #N/A and #REF! from dataset**

```{r}
stock[stock=="#N/A" | stock=="#REF!" ] <- "0"
```


**Convert High and Low stock price columns to numeric values**

```{r}
stock$High.Price <- as.numeric(replace(stock$High.Price,stock$High.Price=="#N/A","0"))
stock$Low.Price <- as.numeric(replace(stock$Low.Price,stock$Low.Price=="#N/A","0"))
stock$Sector <- replace(stock$Sector,stock$Sector%in% c("#N/A","#REF!","0"),"No sector")
tibble(stock)
```

### *High Stock Price mean per Sector*

```{r warning=FALSE}
stock_avg <- stock %>%
  group_by(Sector)%>%
  summarise(HighPriceAvg=round(mean(High.Price,na.rm=TRUE),2),
            LowPriceAvg=round(mean(Low.Price,na.rm=TRUE),2))%>%arrange(desc(HighPriceAvg))#%>%
  #filter(Sector == "No sector")
stock_avg
```

### *Show High stock Price Mean Distribution*

```{r warning=FALSE}
stock_avg%>%ggplot(aes(HighPriceAvg))+geom_histogram(binwidth=1500)+
    labs(x="High Price (Rs.)",
       title="The distribution of low stock price of indian stock is skewed to the right,\n and the data distribution contains meaningless Rs.0 low stock price \n and extreme outliers",
       caption="Data source: https://www.kaggle.com/souravroy1/stock-market-data")+
  theme_bw()

#Remove Outlier for High Price Average
high_no_outlier <-stock_avg%>%
  filter(HighPriceAvg != 0 & HighPriceAvg<9000)

ggplot(high_no_outlier, aes(x=HighPriceAvg))+geom_histogram(aes(y=..density..),binwidth=181)+geom_density(col = "tomato")+
  labs(x="High Price (Rs.)",
       title="The distribution of High stock price of indian stock is bimodal and skewed to the \n right, centered at about Rs.268 with most of the data between Rs.100 and\n Rs.250, a range of roughly Rs.1264 and a few outliers are present on the \n higher end.",
       caption="Data source: https://www.kaggle.com/souravroy1/stock-market-data")+
  theme_bw()

#Summary statistics for high price
summarise(high_no_outlier,mean=mean(HighPriceAvg,na.rm=TRUE),
          median=median(HighPriceAvg,na.rm=TRUE),
          sd=sd(HighPriceAvg,na.rm=TRUE),
          min=min(HighPriceAvg),
          max=max(HighPriceAvg)
          )%>%mutate(Range=max-min)
```



### *Show Low stock Price Mean Distribution*

```{r warning=FALSE}
stock_avg%>%ggplot(aes(LowPriceAvg))+geom_histogram(binwidth=1500,fill="#DAF7A6")+
   labs(x="Low Price (Rs.)",
       title="The distribution of low stock price of indian stock is skewed to the right,\n and the data distribution contains meaningless Rs.0 low stock price \n and extreme outliers",
       caption="Data source: https://www.kaggle.com/souravroy1/stock-market-data")+
  theme_bw()
```


```{r warning=FALSE}
#Remove Outlier from Lower Price Average
low_no_outlier <- stock_avg%>%
  filter(LowPriceAvg !=0 & LowPriceAvg<9000)


ggplot(low_no_outlier,aes(LowPriceAvg))+geom_histogram(aes(y=..density..),binwidth=181,fill="#DAF7A6")+geom_density(col = "red")+
 labs(x="Low Price (Rs.)",
       title="The distribution of low stock price of indian stock is bimodal and skewed to the\n right, centered at about Rs.262 with most of the data between Rs.100 and\n Rs.250, a range of roughly Rs.1239 and a few outliers are present on the \n higher end.",
       caption="Data source: https://www.kaggle.com/souravroy1/stock-market-data")+
  theme_bw()
 


#Summary statistics for Low price Average
summarise(low_no_outlier,mean=mean(LowPriceAvg,na.rm=TRUE),
          median=median(LowPriceAvg,na.rm=TRUE),
          sd=sd(LowPriceAvg,na.rm=TRUE),
          min=min(LowPriceAvg),
          max=max(LowPriceAvg)
          )%>%mutate(Range=max-min)

```









