---
title: "World Annual GDP"
author: "Nnaemeka Okereafor"
date: "10/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
library(tidyverse)
```

### About The Project

The first 5 countries with top GDP and the last 5 countries with least GDP. Also, the prediction of average GDP of the years were xrayed. 

### Read Data from Github

```{r}
df <- read.csv(file="https://raw.githubusercontent.com/nnaemeka-git/global-datasets/main/GDP%20dataset.csv", sep=",",skip=4)
glimpse(df)
```

**Transform dataset with pivot longer**

```{r}
df_long <- df %>% 
  pivot_longer(!c("Country.Name", "Country.Code", "Indicator.Name", "Indicator.Code"),names_to="Year",values_to="GDP")

head(df_long,n=15)
```

**Get Year**

```{r}
df_long$Year <- as.numeric(unlist(str_match_all(df_long$Year,"\\d+..")))
```

**Top 5 GDP**

```{r}
df_sel_val <- df_long %>% select(Country.Name,Country.Code,Year,GDP) %>%
  group_by(Country.Name)%>%
  summarise(Avg_GDP=round(mean(GDP,na.rm=TRUE),2),Count=sum(!is.na(GDP)))%>%arrange(desc(Avg_GDP))%>%
  filter(is.na(Avg_GDP) == FALSE)

#Top 5
top_5_gdp <- df_sel_val[1:5,]
top_5_gdp

top_5_gdp%>% ggplot(aes(reorder(Country.Name,Avg_GDP),Avg_GDP))+
  geom_col(fill="#8CD71A")+geom_text(aes(label=Avg_GDP),color="blue")+ labs(x="Country (billion $)",y="GDP (billion $)", title="Equatorial Guinea has the highest average GDP of 13.67 billion dollar followed by Bosnia\n and Herzegovina with a GDP of 8.97 billion dollar")
```

**Least 5 GDP**

```{r}

least_5_gdp <- tail(df_sel_val,n=5)
least_5_gdp

least_5_gdp%>% ggplot(aes(reorder(Country.Name,Avg_GDP),Avg_GDP))+
  geom_col(fill="#D77E1A")+geom_text(aes(label=Avg_GDP),color="blue")+ labs(x="Country",y="GDP (billion $)", title="South Sudan has the least average GDP with with a negative balance\n of approximately 4.9 billion. Northern Mariana Islands is the second\n least in the list of 5 lowest average GDP")

```

**World Average GDP Growth by Year**

```{r message=FALSE, warning=FALSE, echo=FALSE}
df_sel_val2 <- df_long %>% select(Country.Name,Country.Code,Year,GDP) %>%
  group_by(Year)%>%
  summarise(Avg_GDP=round(mean(GDP,na.rm=TRUE),2),CounOfGDP=sum(!is.na(GDP)))%>%arrange(desc(Avg_GDP))%>%
  filter(is.na(Avg_GDP) == FALSE)

df_sel_val2

df_sel_val2 %>%ggplot(aes(x=Year,y=Avg_GDP))+
  geom_line()+geom_text(aes(label=Avg_GDP),color="blue")+
  labs(y="GDP (billion $)", title="Year 2020 was has the lowest average annual GDP while year 1970 delivered\n the highest average annual GDP")
``` 

### Prediction Analysis

```{r}
#Plot
plot(df_sel_val2$Year,df_sel_val2$Avg_GDP)


#Correlation between year and average GDP
cor(df_sel_val2$Year,df_sel_val2$Avg_GDP)

#Save model in an object
result <- lm(Avg_GDP~Year,data=df_sel_val2)

#Draw regression line
abline(result)

#Display statistics
summary(result)

#Display intercept and Slope
result

#To know the contents of result
#names(result)

#View fitted values
#result$fitted

#Plot year against fitted values
plot(df_sel_val2$Year,result$fitted)

```

**Create function to predict global GDP**

```{r}
Global_GDP <-function(year){
  GDP <- vector()
  for (i in 1:length(year)){
    GDP[i] <-  (result$coef[2]*year[i])+result$coef[1]
  }
  return(GDP)
 }
```

**Make Prediction**

```{r}
#Predict year 1970, 1969 and 2020 using created function above
Global_GDP(c(1970,1969,2020))

#Predict year 1970, 1969 and 2020 using regression default function
predict(result,data.frame(Year=c(1970,1969,2020)))

#Make prediction for the years in the dataset
df_sel_val2$Predicted <-Global_GDP(df_sel_val2$Year)
df_sel_pred <- mutate(df_sel_val2,
                      PercentDifference=((abs((df_sel_val2$Avg_GDP-df_sel_val2$Predicted))/df_sel_val2$Predicted))*100)

head(df_sel_pred)
```

**To compare Year 2020 actual and predicted average GDP**

```{r}
#View the actual and predicted Average GDP for year 2020
pred_2020 <- df_sel_pred%>% filter(Year==2020)
pred_2020
```









